---
title: "pee_poop_times"
author: "Justin Rose"
date: "1/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message=F, warning=F)
library(tidyverse)
library(janitor)
library(lubridate)
library(zoo)
```

We recently got a new golden retriever puppy! In our quest to avoid any accidents inside the house (and the product of having way too much time on our hands due to COVID), I decided to track when our little Lambeau went pee and poop over the span of a month to see if I could try and mitigate the amount of indoor accidents. 

I started tracking this by setting up a routine through my iPhone which would populate a Google Sheet with the exact time and type of potty break that occured. 

Here's some links on how I set up the data collection aspect of this!

```{r}
pp_dta <- read_csv('pee_poop_times.csv')
```


```{r echo=FALSE}
 pee_hist <- pp_dta %>% mutate(test = str_remove(time," EST")) %>% 
  separate(test, c("date","time_clean"),sep = ",") %>% filter(type == 'pee') %>% 
  mutate(date = mdy(date),
         time_clean = as.POSIXct(time_clean, format = "%I:%M:%S %p")) %>% 
  mutate(time_rounded = round_date(time_clean, "30 min")) %>% filter(!is.na(time_rounded)) %>% 
  ggplot(aes(x = time_rounded, fill = inside)) + geom_histogram(binwidth = 60*30) + scale_y_continuous(breaks = seq(0,10,2)) +
  scale_x_datetime(breaks = "1 hour", date_labels = "%H:%M") + theme(axis.text.x = element_text(angle = 45), legend.position = 'bottom')  + scale_fill_manual(values = c("#FF0000", "#d5b60a")) +
  labs(y="# of Times",
       title="Pee Times of Lambeau",
       x="Time of Incident",
       fille="Accident?")

 poophist <- pp_dta %>% mutate(test = str_remove(time," EST")) %>% 
  separate(test, c("date","time_clean"),sep = ",") %>% filter(type == 'poop') %>% 
  mutate(date = mdy(date),
         time_clean = as.POSIXct(time_clean, format = "%I:%M:%S %p")) %>% 
  mutate(time_rounded = round_date(time_clean, "30 min")) %>% filter(!is.na(time_rounded)) %>% 
  ggplot(aes(x = time_rounded, fill = inside)) + geom_histogram(binwidth = 60*30)  + scale_y_continuous(breaks = seq(0,10,2)) +
  scale_x_datetime(breaks = "30 min", date_labels = "%H:%M") + theme(axis.text.x = element_text(angle = 45), legend.position = 'bottom') + scale_fill_manual(values = c("#FF0000","#7b5c00"))+
  labs(y="# of Times",
       title="Poop Times of Lambeau",
       x="Time of Incident",
       fille="Accident?")

 library(plotly)
 pee_hist
 
```



```{r}
 	poophist
```
